---
title: "Week 8: Logistic Regression"
output: html_notebook
---

The Smarket dataset contains daily percentage returns for the S&P 500 stock index over a period of 5 years (2001 to 2005). It includes variables such as:

Year: The year of the observation.
Lag1 to Lag5: Percentage returns for the previous five trading days.
Volume: The volume of shares traded on the given day (in billions).
Today: The percentage return for the current day.
Direction: A categorical variable indicating whether the market went up or down on the current day ("Up" or "Down").


A percentage return is a measure of the profit or loss of an investment relative to its initial value, expressed as a percentage. It shows how much the value of an asset (like a stock or index) has increased or decreased over a specific period, relative to its starting value.


```{r}
install.packages("ISLR")
library(ISLR)
summary(Smarket)
```

```{r}
str(Smarket)
```


```{r}
library(psych)
pairs.panels(Smarket[-9])
```

```{r}
attach(Smarket)
par(mfcol=c(2,3))   #this line creates a 2*3 plotting window
plot(Lag1 ~ Direction,main="lag1", col="red")
plot(Lag2 ~ Direction,main="lag2", col="red")
plot(Lag3 ~ Direction,main="lag3", col="red")
plot(Lag4 ~ Direction,main="lag4", col="red")
plot(Lag5 ~ Direction,main="lag5", col="red")
plot(Volume ~ Direction,main="Volume", col="red")
```

since the observations have temporal order, it is not appropriate to do a random split; instead, we use earlier observations to predict future observations
```{r}
inTrain=Year<2005  # returns a Boolean vector
Smarket.train= Smarket[inTrain,]
Smarket.test= Smarket[!inTrain,]
dim(Smarket.train)
```

```{r}
dim(Smarket.test)
```

```{r}
logistic.m = glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Volume, data=Smarket.train, family = "binomial")
summary(logistic.m)
```
```{r}
preds = predict(logistic.m, Smarket.test, type="response")
head(preds)
```

```{r}
predicted.label=factor(ifelse(preds>0.5, "Up","Down"))
t = table(predicted.label, Smarket.test$Direction)
t
```

The test error of this model is 51.9% which is makes the model practically useless and worse than a random guess. This is disappointing but not surprising because one would generally not expect to predict the next day return from the previous days’ returns. 
```{r}
(t[1,2]+t[2,1])/sum(t)
```

You can check which level is positive by using $levels(your\_response\_variable)$, the positive class is typically the second level in your response factor. For example, if the output is Output: [1] ``Down'' ``Up'', then ``Up'' is positive

```{r}
levels(Smarket$Direction)
```
